---
title: "CMC graphs"
author: "Maurice Richard"
date: "27 januari 2018"
output: html_document
---

```{r setup, include = FALSE}

    knitr::opts_chunk$set(echo = TRUE)
    options(scipen=999)
    
```

```{r libraries, include = FALSE, warning = FALSE}

    #detach("package:plyr", unload=TRUE)
    library(tidyr)      # data wrangling
    library(dplyr)      # data wrangling
    library(lubridate)  # dates
    library(magrittr)   # pipelines
    library(knitr)      # kable
    library(ggplot2)    # plotting
    library(quantmod)

```

```{r Load data, cache = FALSE}

    df_marketdata <- read.csv(file = "Crypto-Markets.csv", stringsAsFactors = FALSE, sep = ";")

    # Change data types
    df_marketdata$date %<>% as.Date("%d-%m-%Y") 

```

We need to clean the data to make sure we do not start with bad data. The first step is to check whether there are any NA's or strange values in the data that we obtained from CMC. 

```{r Check for data quality}

    marketdata_NA <- df_marketdata %>%
        select_if(function(x) any(is.na(x))) %>% colnames()
    
    print(marketdata_NA[-1])
    print(su)
    any(df_marketdata, is.na(date))
    
```

Some coins have duplicate rows, for now we exclude these completely.
```{r Cleaning duplicates}

    # Check for duplicate rows in the data
    duplicates <- df_marketdata %>% group_by(name, date) %>% summarise(count = n()) %>% filter(count != 1)
    df_marketdata %<>% filter(!name %in% duplicates$name)
    
    # excluded the following coins due to duplicates
    print(table(duplicates$name))

```

Some coins might have missing values. We check for each coin whether it is missing values after the first observed value.

```{r Missing values at end of timeseries}

    # first and last are reversed because dataset is sorted descending wise for dates
    coins <- df_marketdata %>% group_by(name) %>% summarise(daysdata = n(), first_date = last(date), last_date = first(date))
    View(coins[coins$last_date != max(coins$last_date), ])
    View(table(df_marketdata$date))
    coins$days <- coins$last_date - coins$first_date + 1
    coins$missing <- coins$days - coins$daysdata
    coins %<>% filter(missing != 0)
    
    missing <- coins[coins$last_date != max(coins$last_date), c(name, )]
    View(df_marketdata %>% filter(name %in% missing))  

```


    
```{r Data preparation}
 


    test <- df_marketdata %>% filter(name %in% | y>4))


    # Calculate rank for each day in the past. 
    df_marketdata        %<>% 
        arrange(ranknow, date)  %>% 
        group_by(ranknow) %>%
        mutate(prevclose = dplyr::lag(close, n = 1, default = NA)) %>%
        mutate(percchg = (close - prevclose) / prevclose * 100) %>%
        arrange(ranknow, desc(date))

    df_marketdata %<>%
        group_by(date) %>% 
        mutate(rank_then = dense_rank(desc(market)))
    
    
```

With the CMC data complete and checked we can analyse the market and different crypto coins.


```{r Rank on date}

    SDate <- "2017/10/01"
    EDate <- "2017/11/30"
    hist_line   <- df_marketdata %>% filter(symbol == "FUN")
    pred_period <- hist_line %>% filter(date >= SDate & date <= EDate)
    
    linearMod   <- lm(rank_then ~ date, data = pred_period)
    rankPred    <- predict(linearMod, pred_period)
    pred_line   <- data.frame(date = seq(from = as.Date(EDate), 
                                      to = as.Date(SDate), 
                                      by = -1),
                              rank_then = rankPred)

    ggplot()+ 
      geom_line(data = hist_line, aes(x = date, y = rank_then), colour="black") +
      geom_line(data = pred_line,  aes(x = date, y = rank_then), colour="red")+
      scale_y_continuous(limits = c(0, 150))+
      scale_x_date(date_breaks = '1 month')+
      ggtitle("Historical rank")+
      xlab("Date")+
      ylab("Rank")

   # modelSummary <- summary(linearMod)  # capture model summary as an object
   # modelCoeffs <- modelSummary$coefficients  # model coefficients



```

The below table shows the number of days a coin was present in the top 5 since the start of 2017.

```{r Number of coins that reached top 5 in 2017}

  sort(table(df_marketdata[df_marketdata$rank_then <= 5 & df_marketdata$date >= "2017/01/01", ]$symbol), decreasing = TRUE)

```

```{r Gain of top coins}

  top_coins <- c("BTC", "ETH", "LTC", "XMR", "DASH")
  SDate <- "2017-11-01"
  EDate <- max(df_marketdata$date, na.rm = TRUE)
  df_top5 <- df_marketdata %>% 
      select(symbol, date, market) %>% 
      filter(symbol %in% top_coins & (date == SDate | date == EDate)) %>% 
      group_by(symbol) %>% 
      spread(date, market)
  
  colnames(df_top5) <- c("symbol", "start", "end")
  
  df_top5$gain <- (df_top5$end / df_top5$start - 1) * 100
  
  print(df_top5)
  
```


```{r Gain of all coins}

  SDate <- "2017-10-01"
  EDate <- max(df_marketdata$date, na.rm = TRUE)
  
  coins <- df_marketdata                    %>% 
      filter(date == SDate | date == EDate) %>% 
      group_by(symbol)                      %>% 
      summarize(count = n())                %>% 
      filter(count == 2)                    %>%
      select(symbol)                        

  df_all_gains <- df_marketdata %>% 
      select(symbol, date, market) %>% 
      filter(symbol %in% coins$symbol & (date == SDate | date == EDate)) %>% 
      group_by(symbol) %>% 
      spread(date, market)
  
  colnames(df_all_gains) <- c("symbol", "start", "end")
  df_all_gains %<>% filter(start > 0 & end > 0)
  
  df_all_gains$gain <- (df_all_gains$end / df_all_gains$start - 1) * 100
  df_all_gains <- df_all_gains[order(df_all_gains$start, decreasing = TRUE), ]
  
  print(df_all_gains)
  
  
```

```{r Returns vs Market cap}

     ggplot(df_all_gains[50:100, ],
        aes(x = log(start), y = gain)) + 
        geom_point(shape=1)+
        #geom_text(aes(label=symbol),vjust=1)+
        scale_y_continuous(limits = c(-50, 400))+
        ggtitle("")+
        xlab("Log(Marketcap)")+
        ylab("Gain")

    mean(df_all_gains$gain[1:10])
    mean(df_all_gains$gain[11:20])
    mean(df_all_gains$gain[21:30])
    mean(df_all_gains$gain[31:40])
    mean(df_all_gains$gain[41:50])
    mean(df_all_gains$gain[51:60])
    mean(df_all_gains$gain[61:70])
    mean(df_all_gains$gain[71:80])
    mean(df_all_gains$gain[81:90])
    mean(df_all_gains$gain[91:100])
    
    mean(df_all_gains$gain[1:20])
    mean(df_all_gains$gain[21:40])
    mean(df_all_gains$gain[41:60])
    mean(df_all_gains$gain[61:80])
    mean(df_all_gains$gain[81:100])


```

```{r Max gains}

    df_max_gains <- df_marketdata %>% 
        select(symbol, date, market) %>% 
        filter(symbol %in% coins$symbol & (date >= SDate)) %>%
        group_by(symbol) %>%
        summarize(start = last(market), max = max(market, na.rm = TRUE))
    
    df_max_gains %<>% filter(start > 0 & max > 0)
    
    df_max_gains$gain <- (df_max_gains$max / df_max_gains$start - 1) * 100
    df_max_gains <- df_max_gains[order(df_max_gains$start, decreasing = TRUE), ]
    
    print(df_max_gains)
  
    mean(df_max_gains$gain[1:20])
    mean(df_max_gains$gain[21:40])
    mean(df_max_gains$gain[41:60])
    mean(df_max_gains$gain[61:80])
    mean(df_max_gains$gain[81:100])
  
```




```{r Select Coin, include = FALSE}

    BTC <- df_marketdata %>% 
           filter(symbol == "MONA")
    ETH <- df_marketdata %>% 
           filter(symbol == "ETH")

```


```{r Regular plot, out.width = '200%'}
    
    ggplot(BTC,
        aes(x = date, y = close)) + 
        geom_bar(stat = "identity")+
        scale_x_discrete(breaks = seq(0, 3000, 100))+
        scale_y_continuous(breaks = seq(0, 20000, 1000))+
        ggtitle("Historical BTC price")+
        xlab("Date")+
        ylab("Price in USD")

```


```{r Candle plot, out.width = '200%'}

    BTC <- BTC[1:150, ]
    BTC$date %<>% as.POSIXct(tz = "")  
    BTC$chg <- ifelse(BTC$close > BTC$open, "up", "dn")
    BTC$width <- as.numeric(periodicity(BTC$date)[1])
    BTC$flat_bar <- BTC[, "high"] == BTC[, "low"]
        
# Candle chart:
    pl <- ggplot(BTC, aes(x = date))+
      geom_linerange(aes(ymin = low, ymax = high)) +
      theme_bw() +
      ggtitle("BTC")+
      xlab("Date")+
      ylab("Price in $")+
      geom_rect(aes(xmin = date - width/2 * 0.9, xmax = date + width/2 * 0.9, 
                    ymin = pmin(open, close), ymax = pmax(open, close), fill = chg)) +
      guides(fill = FALSE, colour = FALSE) + 
      scale_fill_manual(values = c("dn" = "darkred", "up" = "darkgreen"))+
      scale_y_continuous(breaks = seq(0, 20000, 1000))
    
  # Handle special case of drawing a flat bar where OHLC = Open:
    if (any(BTC$flat_bar)) pl <- pl + geom_segment(data = BTC[BTC$flat_bar,], aes(x = date - width / 2 * 0.9, y = close, yend = close, xend = date + width / 2 * 0.9))
    
    print(pl)
       
```
 
 **Question: How do the different crypto's correlate with each other?**
 
```{r Correlations}

    BTC <- df_marketdata %>% 
           filter(name == "Bitcoin")
    ETH <- df_marketdata %>% 
           filter(name == "IOTA")

    BTC <- BTC[1:156, ]
    ETH <- ETH[1:156, ]

    ETH %<>% 
        arrange(ranknow, date)  %>% 
        mutate(prevclose = dplyr::lag(close, n = 1, default = NA)) %>%
        mutate(percchg = (close - prevclose) / prevclose * 100) %>%
        arrange(ranknow, desc(date))
    
    ETH$percchg<- ifelse(is.na(ETH$percchg), 0, ETH$percchg) 
    scatter <- bind_rows(ETH[ ,c("date", "name", "percchg")], BTC[,c("date", "name", "percchg")])
    scatter2 <- scatter %>% spread(key = name, value = percchg)
    cor(scatter2$Bitcoin, scatter2$IOTA)

```


```{r Scatterplot}
     ggplot(scatter,
        aes(x = date, y = percchg)) + 
        geom_bar(aes(fill = name), stat = "identity", position = "dodge")+
        scale_x_discrete(breaks = seq(0, 3000, 100))+
        scale_y_continuous(breaks = seq(0, 20000, 1000))+
        ggtitle("Daily percentage change in close ")+
        xlab("Date")+
        ylab("Daily % change")
```

 